\LARGE \bf Efficient Sampling for Grasp Quality Evaluation on Gaussian Process Implicit Surface Representations [v-12 09-30-2014]

### 2014-10-01 paper-pieter-feedback.tex
mostly merge some pieter edits

%%PA: this citation type-setting is a bit funny to me, I am much more used to [12, 8, 10] or [12], [8], [10] rather than the current paper's typesetting as [12] [8] [10].  If you have seen it in a reputable place and it's the style you guys prefer then I am fine with it, of course
%%ZM: will fix as going through

\todo{Get High res GPIS visualizations, Incorporate next round of feedback}
Consider a robot packing boxes in a shipping warehouse environment. The robot will frequently encounter objects that it has not seen before. The robot will have to plan stable grasps without any prior information other than its sensor data. As seen in Fig. \ref{fig:noisy data}, with today's RGB and depth sensors, surface properties, such as specularity and transparency, can greatly effect the returned observations. Grasp planners based on deterministic algorithms are not designed to handle such noise and could fail. To plan grasps in a partially observed world, recent research has focused on incorporating uncertainity%%PA: --> spelling
 in the shape of the object \cite{kehoe2012estimating}, \cite{christopoulos2007handling}, \cite{zheng2005}. %%PA: why two dots? %%PA: and same paper is cited twice... %% ZM fixed

While past work has encoded uncertainty in a polygonal mesh representation \cite{kehoe2012estimating}, an alternative is to work with the signed distance fields (SDFs), a continuous valued function that is an implicit surface: zero-valued at the surface of the object, positive value outside the surface and negative valued inside the surface. This representation is growing in popularity in the robotics field for its ability to integrate range data in real time \cite{newcombe2011kinectfusion}\cite{curless1996volumetric}. To incorporate uncertainty one can use a Gaussian Process Implicit Surface (GPIS), a Gaussian process over signed distance field. Not only does GPIS present a straightforward and formal way to incorporate measurements from visual, laser and tactile sensors, they  also allow for a continuous differentiable function on shape, 
%%PA: continuous and differentiable (which I believe would be a redundant statement as differentiable implies continuous), or continuously differentiable?
which can be used for a feedback grasp controller, finding locally optimal grasps and active exploration  \cite{dragiev2011}, \cite{jeffs}, \cite{hollinger2013}.

>>>>

%%PA: this citation type-setting is a bit funny to me, I am much more used to [12, 8, 10] or [12], [8], [10] rather than the current paper's typesetting as [12] [8] [10].  If you have seen it in a reputable place and it's the style you guys prefer then I am fine with it, of course
%%ZM: fixed

\todo{Get High res GPIS visualizations, Incorporate next round of feedback}
Consider a robot packing boxes in a shipping warehouse environment. The robot will frequently encounter objects that it has not seen before. The robot will have to plan stable grasps without any prior information other than its sensor data. As seen in Fig. \ref{fig:noisy data}, with today's RGB and depth sensors, surface properties, such as specularity and transparency, can greatly effect the returned observations. Grasp planners based on deterministic algorithms are not designed to handle such noise and could fail. To plan grasps in a partially observed world, recent research has focused on incorporating uncertainity%%PA: --> spelling
 in the shape of the object \cite{kehoe2012estimating}, \cite{christopoulos2007handling}, \cite{zheng2005}. %%PA: why two dots? %%PA: and same paper is cited twice... %%ZM: fixed

While past work has encoded uncertainty in a polygonal mesh representation \cite{kehoe2012estimating}, an alternative is to work with the signed distance fields (SDFs), a continuous real-valued function that is an implicit surface: zero-valued at the surface of the object, positive value outside the surface and negative valued inside the surface. This representation is growing in popularity in the robotics field for its ability to integrate range data in real time \cite{newcombe2011kinectfusion}, \cite{curless1996volumetric}. To incorporate uncertainty one can use a Gaussian Process Implicit Surface (GPIS), a Gaussian process over signed distance field. Not only does GPIS present a straightforward and formal way to incorporate measurements from visual, laser and tactile sensors, they also allow for a differentiable function as a shape representation, 
%%PA: continuous and differentiable (which I believe would be a redundant statement as differentiable implies continuous), or continuously differentiable? %%ZM: fixed
which can be used for a feedback grasp controller, finding locally optimal grasps and active exploration  \cite{dragiev2011}, \cite{jeffs}, \cite{hollinger2013}.

### 2014-10-01 paper.tex
mostly merge some pieter edits

\todo{Get High res GPIS visualizations, Incorporate next round of feedback}
Consider a robot packing boxes in a shipping warehouse environment. The robot will frequently encounter objects that it has not seen before. The robot will have to plan stable grasps without any prior information other than its sensor data. As seen in Fig. \ref{fig:noisy data}, with today's RGB and depth sensors, surface properties, such as specularity and transparency, can greatly effect the returned observations. Grasp planners based on deterministic algorithms are not designed to handle such noise and could fail. To plan grasps in a partially observed world, recent research has focused on incorporating uncertainity in the shape of the object \cite{kehoe2012estimating} \cite{kehoe2012estimating} \cite{christopoulos2007handling} \cite{zheng2005}..

While past work has encoded uncertainty in a polygonal mesh representation \cite{kehoe2012estimating}, an alternative is to work with the signed distance fields (SDFs), a continuous valued function that is an implicit surface: zero-valued at the surface of the object, positive value outside the surface and negative valued inside the surface. This representation is growing in popularity in the robotics field for its ability to integrate range data in real time \cite{newcombe2011kinectfusion}\cite{curless1996volumetric}. To incorporate uncertainty one can use a Gaussian Process Implicit Surface (GPIS), a Gaussian process over signed distance field. Not only does GPIS present a straightforward and formal way to incorporate measurements from visual, laser and tactile sensors, they  also allow for a continuous differentiable function on shape, which can be used for a feedback grasp controller, finding locally optimal grasps and active exploration  \cite{dragiev2011}\cite{jeffs}\cite{hollinger2013}.

\begin{figure}[ht!]
\centering
\includegraphics[width=8.5cm,height=4cm]{figures/Slide02.jpg}
\caption{\footnotesize
Left: HD image of a common measuring cup from in a household environment. Right: Image taken from PR2's Primesense Carmine sensor, large parts of the measuring cup are not detected by the depth sensor and are not visible in the RGB image.}
\vspace*{-10pt}
\label{fig:noisy data}
\end{figure}

Given a method to represent shape uncertainty, a common technique to evaluate grasp quality is to sample from the distribution on shape and perform Monte-Carlo integration on a given grasp metric \cite{kehoe2012estimating}\cite{kehoe2012toward}\cite{christopoulos2007handling}. Applying this technique to GPIS requires an exhaustive sampling of the signed distance field, which scales $O(n^6)$ for a 2D $n \times n$ workspace and $O(n^9)$ for a 3D workspace.The computationally complexity of sampling is derived from the fact sampling from a Gaussian requires matrix inversion of the entire signed distance field. Our paper reduces this complexity to $O(n^3)$ by using the insight that for grasp quality evaluation one only cares about the parameters that define a grasp: contacts points, surface normals, center of mass and friction coefficient \cite{pokorny2013classical}. We constrain the contact points by considering the approach trajectory of a gripper.  Hence, sampling the whole signed distance field is not needed, but instead one can only work with the distributions on grasp parameters given the shape uncertainty model and gripper trajectory. In light of this our contributions are the following:

\begin{enumerate}
	\item Derivation of  distributions on grasp parameters given GPIS shape uncertainty model and approach trajectory of gripper.
	\item  Demonstrating both theoretically and empirically how those distributions reduce the computationally complexity of sampling.
	\item Applying adaptive sampling techniques to efficiently find high quality grasps in a given set of proposed grasps. 
\end{enumerate}

@@

Where $Q(g|S,\Gamma)$ is the grasp quality that is computed on a shape sample drawn from $p(S)$. To compute this we intersect the zero crossing of the level set with the propose grasp plan $\Gamma$ and determine the parameters $g$, this has been the approach taken in previous work  \cite{kehoe2012estimating} \cite{kehoe2012toward} \cite{christopoulos2007handling}. See Fig. \ref{fig:shape_samples} for an example of what samples drawn from $p(S)$ induced by GPIS look like. For computational reasons we approximate the integral via Monte-Carlo Integration. We use importance sampling to draw from the distribution induced by GPIS and calculate the following

>>>>

%%PA: this citation type-setting is a bit funny to me, I am much more used to [12, 8, 10] or [12], [8], [10] rather than the current paper's typesetting as [12] [8] [10].  If you have seen it in a reputable place and it's the style you guys prefer then I am fine with it, of course
%%ZM: fixed

\todo{Get High res GPIS visualizations, Incorporate next round of feedback}
Consider a robot packing boxes in a shipping warehouse environment. The robot will frequently encounter objects that it has not seen before. The robot will have to plan stable grasps without any prior information other than its sensor data. As seen in Fig. \ref{fig:noisy data}, with today's RGB and depth sensors, surface properties, such as specularity and transparency, can greatly effect the returned observations. Grasp planners based on deterministic algorithms are not designed to handle such noise and could fail. To plan grasps in a partially observed world, recent research has focused on incorporating uncertainity%%PA: --> spelling
 in the shape of the object \cite{kehoe2012estimating}, \cite{christopoulos2007handling}, \cite{zheng2005}. %%PA: why two dots? %%PA: and same paper is cited twice... %%ZM: fixed

While past work has encoded uncertainty in a polygonal mesh representation \cite{kehoe2012estimating}, an alternative is to work with the signed distance fields (SDFs), a continuous real-valued function that is an implicit surface: zero-valued at the surface of the object, positive value outside the surface and negative valued inside the surface. This representation is growing in popularity in the robotics field for its ability to integrate range data in real time \cite{newcombe2011kinectfusion}, \cite{curless1996volumetric}. To incorporate uncertainty one can use a Gaussian Process Implicit Surface (GPIS), a Gaussian process over signed distance field. Not only does GPIS present a straightforward and formal way to incorporate measurements from visual, laser and tactile sensors, they also allow for a differentiable function as a shape representation, 
%%PA: continuous and differentiable (which I believe would be a redundant statement as differentiable implies continuous), or continuously differentiable? %%ZM: fixed
which can be used for a feedback grasp controller, finding locally optimal grasps and active exploration  \cite{dragiev2011}, \cite{jeffs}, \cite{hollinger2013}.

\begin{figure}[ht!]
\centering
\includegraphics[width=8.5cm,height=4cm]{figures/Slide02.jpg}
\caption{Left: HD image of a common measuring cup from in a household environment. Right: Image taken from PR2's Primesense Carmine sensor, large parts of the measuring cup are not detected by the depth sensor and are not visible in the RGB image.}
\vspace*{-10pt}
\label{fig:noisy data}
\end{figure}

Given a method to represent shape uncertainty, a common technique to evaluate grasp quality is to sample from the distribution on shape and perform Monte-Carlo integration on a given grasp metric \cite{kehoe2012estimating}, \cite{kehoe2012toward}, \cite{christopoulos2007handling}. Applying this technique to GPIS requires an exhaustive sampling of the signed distance field, which scales $O(n^6)$ for a 2D $n \times n$ workspace and $O(n^9)$ for a 3D workspace.The computationally complexity of sampling is derived from the fact sampling from a Gaussian requires matrix inversion of the entire signed distance field. Our paper reduces this complexity to $O(n^3)$ by using the insight that for grasp quality evaluation one only cares about the parameters that define a grasp: contacts points, surface normals, center of mass and friction coefficient \cite{pokorny2013classical}. 
%%PA: that's neat!  let's have you include some of that in the abstract!
%%PA: also n^3 independent as to whether it's a 2-D or 3-D problem, or is n^3 specific to a particular dimensionality?
We constrain the contact points by considering the approach trajectory of a gripper.  Hence, sampling the whole signed distance field is not needed, but instead one can only work with the distributions on grasp parameters given the shape uncertainty model and gripper trajectory. In light of this our contributions are the following:

\begin{enumerate}
	\item Derivation of distributions on grasp parameters given GPIS shape uncertainty model and approach trajectory of gripper.%%PA: grammar off
	\item Demonstrating both theoretically and empirically how those distributions reduce the computational complexity of sampling.%%PA: -> computational %%ZM: fixed
	\item Applying adaptive sampling techniques to efficiently find high quality grasps in a given set of proposed grasps. 
\end{enumerate}

@@

Where $Q(g|S,\Gamma)$ is the grasp quality that is computed on a shape sample drawn from $p(S)$. To compute this we intersect the zero crossing of the level set with the propose grasp plan $\Gamma$ and determine the parameters $g$, this has been the approach taken in previous work  \cite{kehoe2012estimating}, \cite{kehoe2012toward}, \cite{christopoulos2007handling}. See Fig. \ref{fig:shape_samples} for an example of what samples drawn from $p(S)$ induced by GPIS look like. For computational reasons we approximate the integral via Monte-Carlo Integration. We use importance sampling to draw from the distribution induced by GPIS and calculate the following
